---
title: "Practical Machine Learning Project"
author: "Shreyash Mishra"
date: "August 6, 2020"
output: html_document
---


## Loading required libraries
```{r}
library(caret)
library(rpart)
library(rpart.plot)
library(randomForest)
library(rattle)
```


## Taking in the data
```{r}
set.seed(1)
training <- read.csv("pml-training.csv")
testing <- read.csv("pml-testing.csv")
```


## Exploring the data
```{r}
dim(training)
dim(testing)
head(training)
```


Checking if the training dataset has any null values,

```{r}
sum(complete.cases(training))
colnames(training)
plot(colMeans(is.na(training)))
```

There are columns with a lot of missing values.  
We will reatain only the columns without NA values  


## Cleaning the data
```{r}
features <- names(testing[,colSums(is.na(testing)) == 0])[8:59]
trainclasse <- training[,c(features,"classe")]
testproblem <- testing[,c(features,"problem_id")]
```


## Partitioning the data

In order to evaluate our model before submitting it for grading, we'll designate a partition of it for validation.

```{r}
inTrain <- createDataPartition(trainclasse$classe, p=0.7, list = FALSE)
myTraining <- trainclasse[inTrain,]
myTesting <- trainclasse[-inTrain,]
```

## Modeling the data


We will fit a model using **Decision Tree** and **Random Forest**  


## Decision Tree Prediction
```{r}
set.seed(12345)
DTmodel <- rpart(classe ~ ., data = myTraining, method = "class")
fancyRpartPlot(DTmodel)
DTpredict <- predict(DTmodel, myTesting, type = "class")
confusionMatrix(DTpredict, myTesting$classe)
```


## Random Forest Prediction
```{r}
RFmodel <- randomForest(classe ~ ., data = myTraining)
RFpredict <- predict(RFmodel, myTesting, type = "class")
confusionMatrix(RFpredict, myTesting$classe)
```


Since the random forest model's accuracy was 99.3%, the out of sample error is 0.007.
We will use the random forest model to submit our predictions.


```{r}
FinalPredict <- predict(RFmodel, testing, type = "class")
FinalPredict
pml_write_files = function(x){
    n = length(x)
    for(i in 1:n){
        filename = paste0("problem_id_",i,".txt")
        write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
    }
}

pml_write_files(FinalPredict)
```

